---
title: "Title of Report"
author: "Author of this masterpiece"
date: "`r Sys.Date()`"
output:
  pdf_document:
  html_document:
    theme: flatly
    highlight: zenburn
  word_document:
---


```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

library(tidyverse)
library(tidyquant)
library(showtext)
library(ggtext)

# load fonts we will use
font_add_google("Montserrat", "Montserrat") # official LBS font
font_add_google("Lato", "Lato")

## Automatically use showtext to render text for future devices
showtext_auto()
```


The following is a list of the data I download from FRED.

- [`PCU32733273`](https://fred.stlouisfed.org/series/PCU32733273), Cement and Concrete Product Manufacturing 
- [`PIORECRUSDM`](https://fred.stlouisfed.org/series/PIORECRUSDM), Global price of Iron Ore
- [`PCOPPUSDM`](https://fred.stlouisfed.org/series/PCOPPUSDM), Global price of Copper
- [`PALUMUSDM`](https://fred.stlouisfed.org/series/PALUMUSDM), Global price of Aluminum
- [`POILBREUSDM`](https://fred.stlouisfed.org/series/POILBREUSDM), Global price of Brent Crude
- [`WPU07`](https://fred.stlouisfed.org/series/WPU07), Producer Price Index by Commodity: Rubber and Plastic Products
- [`WPS061`](https://fred.stlouisfed.org/series/WPS061), Producer Price Index by Commodity: Chemicals and Allied Products: Industrial Chemicals
- [`PCU3253132531`](https://fred.stlouisfed.org/series/PCU3253132531), Producer Price Index by Industry: Fertilizer Manufacturing

```{r}
#| label: get-data-from-FRED
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

indicators <- c("PCU32733273", # Cement and Concrete Product Manufacturing 
           "PIORECRUSDM", # Global price of Iron Ore
           "PCOPPUSDM", # Global price of Copper
           "PALUMUSDM", # Global price of Aluminum
           "POILBREUSDM", #Global price of Brent Crude
           "WPU07", # Producer Price Index by Commodity: Rubber and Plastic Products
           "WPS061", # Producer Price Index by Commodity: Chemicals and Allied Products: Industrial Chemicals
           "PCU3253132531" #Producer Price Index by Industry: Fertilizer Manufacturing
           ) 


my_data <- indicators |> 
  # Fetch economic data for the specified indicators
  tq_get(get = "economic.data",
         from = "2021-01-01",  # Start date
         to = Sys.Date()) |>   # End date (current system date)
  
  # Group the data by symbol (indicator)
  group_by(symbol) |> 
  
  # Calculate the relative change in price compared to the first observation
  mutate(delta = price / first(price)) |> 
  
  # Remove the grouping
  ungroup() |> 
  
  # Create a new variable with more user-friendly names for each indicator
  mutate(
    new_name = case_when(
      symbol == "PCU32733273" ~ "Cement", 
      symbol == "PIORECRUSDM" ~ "Iron Ore",
      symbol == "PCOPPUSDM" ~ "Copper",
      symbol == "PALUMUSDM" ~ "Aluminum",
      symbol == "POILBREUSDM" ~ "Brent Crude",
      symbol == "WPU07" ~ "Rubber",
      symbol == "WPS061" ~ "Industrial Chemicals",
      symbol == "PCU3253132531" ~ "Fertilizer"
    )
  )



```


## Time series graph of indicators

```{r, echo=FALSE}


my_data |> 
  
  # plot the data using ggplot
  ggplot() +
  # Set aesthetic mappings: x-axis is date, y-axis is price, color by symbol
  aes(x = date, y = price, colour = symbol) +
  
  # Add line geometry to create time series lines
  geom_line() +
  
  # Apply a black and white theme
  theme_bw() +
  
  # Create facets (subplots) for each indicator, wrapped into a 2-column layout
  # with free scales for each facet
  facet_wrap(~new_name,
             ncol = 2,
             scales = "free") +
  
  # Remove the legend
  theme(legend.position = "none") +
  
  # Set the plot title, caption, and remove x and y axis labels
  labs(title = "Price Evolution 2021-01-01 to today",
       caption = "Source: FRED",
       x = NULL, y = NULL) +
  
  # Ensure the title is top-left aligned
  theme(plot.title.position = "plot") +
  
  # Set text properties: size 12, Lato font family
  theme(text = element_text(size = 12, family = "Lato"))


```

## Time series graph of cumulative change

```{r, echo=FALSE}


my_data |> 
  ggplot() +
  # Set aesthetic mappings: x-axis is date, y-axis is delta (relative change), color by symbol
  aes(x = date, y = delta, colour = symbol) +
  
  # Add line geometry to create time series lines
  geom_line() +
  
  # Apply a black and white theme
  theme_bw() +
  
  # Create facets (subplots) for each indicator, wrapped into a 2-column layout
  # with free scales for each facet
  facet_wrap(~new_name,
             ncol = 2,
             scales = "free") +
  
  # Remove the legend
  theme(legend.position = "none") +
  
  # Format y-axis labels as percentages
  scale_y_continuous(label = scales::percent) +
  
  # Set the plot title, caption, and remove x and y axis labels
  labs(title = "Cumulative change since 2021-01-01",
       caption = "Source: FRED",
       x = NULL, y = NULL) +  
  
  # Ensure the title is top-left aligned
  theme(plot.title.position = "plot") +
  
  # Set text properties: size 12, Montserrat font family
  theme(text = element_text(size = 12, family = "Montserrat"))



# rebase everything to 100%
my_data |> 
  mutate(new_name = fct_reorder(new_name, -delta)) |> 
  ggplot()+
  aes(x=date, y = delta, colour = new_name)+
  geom_line()+
  theme_bw()+
  # facet_wrap(~new_name,
  #            ncol = 2,
  #            scales="free")+
  # theme(legend.position = "none")+
  scale_y_continuous(label = scales::percent)+
  labs(title= "Cumulative change since 2021-01-01",
       caption = "Source:FRED",
       x=NULL, y = NULL, 
       colour = "Commodity")+  
  theme(plot.title.position = "plot")+ # ensure title is top-left aligned
  theme(text=element_text(size=12, family="Montserrat"))


```

## Scatterplot - Correlation Matrix

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(GGally)


my_data |> 
  select(date, price, new_name) |> 
  pivot_wider(names_from = new_name,
              values_from = price) |> 
  select(-date) |> 
  ggpairs()+
    labs(title= "Scatterplot- Correlation Matrix",
       caption = "Source:FRED",
       x=NULL, y = NULL, 
       colour = "Commodity")+  
  theme(plot.title.position = "plot")+ # ensure title is top-left aligned
  theme(text=element_text(size=9, family="Lato"))

```

Finally, let us save the file so I can open it in PowerBI-- after all, we all make mistakes ;-)

```{r}
write_csv(my_data, "latest_fred_data.csv")
```

