---
title: "TfL Bikes: How did the groups fare?"
author: "Kostis Christodoulou"
date: "`r Sys.Date()`"
output:
  html_document:
    highlight: zenburn
    theme: flatly
---

```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(GGally)
library(here)
library(skimr)
library(janitor)
library(broom)
library(huxtable)
library(lubridate)
library(ggfortify)
library(performance)
library(ggiraph)
library(car)




```


# London Bikes data

We'll use data from http://www.tfl.gov.uk to analyse usage of the London Bike Sharing scheme. The data you had to fit your model was from July 2010 to Jul 2023. This was the **training** set of your data, and the available variables are as follows


```{r, warning= FALSE, message= FALSE, echo=FALSE}

#read the CSV file
bike <- read_csv(here::here("data", "london_bikes.csv")) %>% 
  mutate(weekend = if_else((wday == "Sat" | wday == "Sun"), TRUE, FALSE))

```


```{r, echo=FALSE}
# fix dates using lubridate, and generate new variables for year, month, month_name, day, and day_of _week
bike <- bike %>%   
  mutate(
    year=year(date),
    month = month(date),
    month_name=month(date, label = TRUE),
    day_of_week = wday(date, label = TRUE)) 

# generate new variable season_name to turn seasons from numbers to Winter, Spring, etc
bike <- bike %>%  
  mutate(
    season_name = case_when(
      month_name %in%  c("Dec", "Jan", "Feb")  ~ "Winter",
      month_name %in%  c("Mar", "Apr", "May")  ~ "Spring",
      month_name %in%  c("Jun", "Jul", "Aug")  ~ "Summer",
      month_name %in%  c("Sep", "Oct", "Nov")  ~ "Autumn",
    ),
    season_name = factor(season_name, 
                         levels = c("Winter", "Spring", "Summer", "Autumn"))) %>% 
  mutate(set = "train")

glimpse(bike)  
```

I wanted to test how well groups model fared with unseen data, namely data from Aug 2023 to Jan 2024. This is the **testing** data, shown in red below

```{r, warning= FALSE, message= FALSE, echo=FALSE}
#read the CSV file
new_bike <- read_csv(here::here("bikes_Aug23_Jan24.csv")) %>% 
  mutate(
    year=year(date),
    month = month(date),
    month_name=month(date, label = TRUE),
    day_of_week = wday(date, label = TRUE)) %>% 
  mutate(
    season_name = case_when(
      month_name %in%  c("Dec", "Jan", "Feb")  ~ "Winter",
      month_name %in%  c("Mar", "Apr", "May")  ~ "Spring",
      month_name %in%  c("Jun", "Jul", "Aug")  ~ "Summer",
      month_name %in%  c("Sep", "Oct", "Nov")  ~ "Autumn",
    ),
    season_name = factor(season_name, 
                         levels = c("Winter", "Spring", "Summer", "Autumn"))) %>% 
  mutate(set = "test")
```


```{r, warning= FALSE, echo=FALSE}

# join both files

bike %>% 
  bind_rows(new_bike) %>% 
  ggplot()+
  aes(x=date, y=bikes_hired, colour = set)+
  geom_point(alpha = 0.3)+
  theme_bw()+
labs(x=NULL, y = NULL, colour = "Data Set", 
     label = "TfL Bike Rentals, Jul 2010- Jan 2024")+
NULL

```



# Model building

The following is a list of models that were submitted by each study group, `model1` corresponds to the best model submitted by Study Group 1, etc.

There are two additional models:

- `model0` is the naive mean model. To predict how many bikes we will rent on a day, we just use the arithmetic mean
- `model99` is a hybrid model, which uses the explanatory variables we have, as well as an auto-regressive AR(1) variable, meaning today's rentals are also explained by what happened yesterday.

```{r}

# Define the models- number corresponds to study group
model1 <- lm(bikes_hired ~ year+ wday + month + cloud_cover + humidity + pressure + 
               radiation + precipitation + sunshine +mean_temp + season_name, data = bike)


model2 <- lm(bikes_hired ~ month_name+day_of_week+humidity+pressure+
               radiation+precipitation+snow_depth+sunshine+
               mean_temp+max_temp, data=bike)

#for model3, I do not use covidTRUE and period, as I dont know how they were defined
# impact on Adj R2 is roughly 1% and residual SE is 5562 instead of 5657 in this definition
model3 <- lm(bikes_hired ~ mean_temp + wday + humidity + pressure + 
               precipitation +season_name+
               sunshine + radiation + snow_depth + month, data = bike)

# Group 4 had the highesrt feature engineering of all. Here is a variant without 
# creating new variables
model4 <- lm(bikes_hired ~ lag(bikes_hired,1) + mean_temp + weekend + season_name + 
               I(year-2009) + humidity+ sqrt(precipitation) + cloud_cover, data = bike)

# unsure how they defined weekday logical
model5 <- lm(bikes_hired ~ mean_temp + sunshine + radiation + weekend + 
               precipitation + snow_depth + humidity, data = bike)

# model6 interactions galore. Explanation?
model6 <- lm(bikes_hired ~ mean_temp*min_temp*max_temp*precipitation*sunshine+
               snow_depth + pressure + wday + year + season_name, data = bike)

model7 <- lm(bikes_hired ~ humidity + year + week + cloud_cover + 
               pressure + precipitation + min_temp + day_of_week, data = bike)


# Let us add a few more models

# Define a model using the mean (average)
model0 <- lm(bikes_hired ~ 1, data = bike)

# define a simpler model with AR(1)
# Auto-regressive of lag 1
model99 <- lm(bikes_hired ~ lag(bikes_hired,1) +
               mean_temp + wday + 
               humidity + precipitation + sunshine + radiation, data = bike) 
```


# Comparison of different models



```{r}

# produce summary table comparing models using huxtable::huxreg()
# produce summary table comparing models using huxtable::huxreg()
huxreg(model1, model2, model3, model4, model5, model6, model7, model0, model99,
       statistics = c('#observations' = 'nobs', 
                      'R squared' = 'r.squared', 
                      'Adj. R Squared' = 'adj.r.squared', 
                      'Residual SE' = 'sigma'), 
       bold_signif = 0.05
) %>% 
  set_caption('Comparison of models')

```


# Collinearity fοr each of the models


```{r}
# Check VIF

vif(model1)
vif(model2)

vif(model3)

vif(model4)

vif(model5)

# because of interaction terms
vif(model6, type = "predictor")

vif(model7)

vif(model99)

```





# Calculating RMSE (Root Mean Squared Error) of Testing Data

The RMSE (Root Mean Squared Error) is the best tool to measure how much we may have overfitted a model. 
We always expect the RMSE in the testing data to increase a bit.

```{r, echo=FALSE, warning = FALSE, message = FALSE}

# model predictions
library(yardstick)
library(kableExtra)

# calcualte RMSE for training data
training1 <- broom::augment(model1) %>% 
  drop_na(.fitted)

training2 <- broom::augment(model2) %>% 
  drop_na(.fitted)

training3 <- broom::augment(model3) %>% 
  drop_na(.fitted)

training4 <- broom::augment(model4) %>% 
  drop_na(.fitted)

training5 <- broom::augment(model5) %>% 
  drop_na(.fitted)

training6 <- broom::augment(model6) %>% 
  drop_na(.fitted)

training7 <- broom::augment(model7) %>% 
  drop_na(.fitted)

training99 <- broom::augment(model99) %>% 
  drop_na(.fitted)

training0 <- broom::augment(model0) %>% 
  drop_na(.fitted)

rmse_train1 <- rmse(training1, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 1, 
         set = "train")


rmse_train2 <- rmse(training2, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 2, 
         set = "train")


rmse_train3 <- rmse(training3, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 3, 
         set = "train")


rmse_train4 <- rmse(training4, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 4, 
         set = "train")


rmse_train5 <- rmse(training5, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 5, 
         set = "train")


rmse_train6 <- rmse(training6, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 6, 
         set = "train")


rmse_train7 <- rmse(training7, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 7, 
         set = "train")


rmse_train0 <- rmse(training0, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 0, 
         set = "train")

rmse_train99 <- rmse(training99, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 99, 
         set = "train")


rmse_train <- 
  rmse_train1 %>% 
  bind_rows(rmse_train2, rmse_train3, rmse_train4, rmse_train5, rmse_train6, rmse_train7, rmse_train0, rmse_train99 ) %>% 
  select(group,  rmse = .estimate, set) 



## Testing datsa

predictions1 <- broom::augment(model1, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions2 <- broom::augment(model2, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions3 <- broom::augment(model3, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions4 <- broom::augment(model4, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions5 <- broom::augment(model5, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions6 <- broom::augment(model6, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions7 <- broom::augment(model7, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions0 <- broom::augment(model0, newdata = new_bike) %>% 
  drop_na(.fitted)

predictions99 <- broom::augment(model99, newdata = new_bike) %>% 
  drop_na(.fitted)



rmse1 <- rmse(predictions1, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 1, 
         set = "test")

rmse2 <- rmse(predictions2, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 2, 
         set = "test")

rmse3 <- rmse(predictions3, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 3, 
         set = "test")

rmse4 <- rmse(predictions4, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 4, 
         set = "test")

rmse5 <- rmse(predictions5, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 5, 
         set = "test")

rmse6 <- rmse(predictions6, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 6, 
         set = "test")

rmse7 <- rmse(predictions7, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 7, 
         set = "test")



rmse0 <- rmse(predictions0, 
              truth = bikes_hired, 
              estimate = .fitted) %>% 
  mutate(group = 0, 
         set = "test")

naive <- rmse0 %>% select(.estimate) %>% pull()

rmse99 <- rmse(predictions99, 
              truth = bikes_hired, 
              estimate = .fitted)  %>% 
  mutate(group = 99, 
         set = "test")


# %>% 
#   mutate(improve_over_mean_rmse = rmse/naive - 1)


rmse_test <- 
  rmse1 %>% 
  bind_rows(rmse2, rmse3, rmse4, rmse5, rmse6, rmse7, rmse0, rmse99)%>% 
  select(group,  rmse = .estimate, set) 



rmse_all_groups <- rmse_train %>%
  bind_rows(rmse_test) %>% 
  pivot_wider(names_from = "set",
              values_from = "rmse") %>% 
  mutate(delta_train_test = test/train-1)

rmse_all_groups %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

# Checking Overfitting

Its firaly easy to keep adding variables and/or feature engineer your model to increase its R2 as much as possible. However, how does your model fare when its applied to unseen data? In our case, we fitter the model in data up to Jul 2023, but we will use data from the lst 6 months (Aug 2023- Jan 2024) to see how well our model fits the data. We calculate the Root Mean Squared Error (RMSE) for both training data (where the model was fitted) and testing data (the unseen data). The following graph shows the difference between training and testing RMSE.

```{r, echo= FALSE}
library(ggrepel)

rmse_all_groups %>% 
  ggplot()+
  aes(y=test, x = train, label=group)+
  geom_point()+
  geom_text_repel(size=4)+
  theme_bw()+
  theme(legend.position = "none")+
  labs(title = "RMSE (Root Mean Squared Error)",
       subtitle = "Numbers refer to study groups, 0 is the naive mean model",
       x = "Train RMSE",
       y = "Test RMSE")+
  xlim(3000,10000)+
  ylim(3000,10000)+
  geom_abline(slope = 1, intercept = 0, color = "#001e62") + # Add a 45-degree line
  NULL
```


